Reasoning about a Capability Machine with Local Capabilities\\ Provably Safe Stack and Return Pointer Management (without OS Support)\\ Technical Appendix Including Proofs and Details

### 2017-09-25 fully-abstract-comp/technical-report.tex
Add a remark and a small fix.

% Rotate entire figures
\usepackage{rotating}

@@

% Four arguments
%\newcommand{\ssubseg}[4]{\fourinstr{\src{subseg}}{#1}{#2}{#3}{#4}}
\newcommand{\ssplit}[4]{\fourinstr{\src{split}}{#1}{#2}{#3}{#4}}

@@

% Four arguments
%\newcommand{\tsubseg}[4]{\fourinstr{\trg{subseg}}{#1}{#2}{#3}{#4}}
\newcommand{\tsplit}[4]{\fourinstr{\trg{split}}{#1}{#2}{#3}{#4}}

@@

\dominique{7-9-2017: Should we keep local/global and linear/non-linear?  I was thinking just the latter would suffice...}
\lau{07-09-2017: The local/global mechanism still has some value as it now can be used for the purpose of arguments that cannot be stored by the receiver (if you are willing to ``pay'' the clearing cost). I do, however, not think that it is strictly necessary here to make things work.}
\lau{12-09-2017: I will remove the local capabilities today for the reasons we have discussed (they add to the complexity and they seem to be of little to no use without a clear memory instruction).}
Source language domains:
\[
  \begin{array}{rrcl}
   \addrbnf,\basebnf \in & \Addr & \defeq & \nats \\
    \sealbasebnf, \sigma \in & \Seal & \defeq & \nats \\
    &\Word & \defeq & \ints \uplus \SealableCaps\\
    \permbnf \in& \Perm & \defbnf & \dots \\
    &\linbnf & \defbnf & \linear \mid \normal \\
    &\aendbnf & \in & \Addr \uplus \{\infty \} \\
    &\sealendbnf & \in & \Seal \uplus \{\infty \} \\
    &\scbnf &\defbnf & ((\permbnf,\linbnf),\basebnf,\aendbnf,\addrbnf) \mid \seal{\linbnf,\sealbasebnf,\sealendbnf,\sigma}\\
    & & & {\sourcecolor{} \mid \stkptr{\permbnf,\basebnf,\aendbnf,\addrbnf}}\\ 
    &\cbnf & \defbnf &  \scbnf \mid \sealed{\scbnf,\sigma}{\sourcecolor{} \; \mid \retptrd(\sigma) \mid \retptrc(\sigma)}\\ 
    &\RegName & \defbnf &\pcreg \mid \rret \mid \rstk \mid \rdata \mid \dots \\
    &\Reg & \defeq & \RegName \fun \Word\\
    &\Mem & \defeq & \Addr \fun \Word \\
    &\MemSeg & \defeq & \Addr \parfun \Word \\
    {\sourcecolor \sstkframe \in} & {\sourcecolor \StkFrame} & {\sourcecolor \defeq} & {\sourcecolor \Caps \times \MemSeg \times \Addr}\\
    \src{\sstk \in}& \src{ \Stack} & \src{ \defeq} & \src{ \StkFrame^*} \\
    \Phi \in & \ExecConf & \defeq & \Mem \times \Reg {\sourcecolor{} \; \times \; \Stack \times \MemSeg }\\
    &\Conf & \defeq & \ExecConf \uplus \{\failed\} \uplus (\{\halted\} \times \Mem)
  \end{array}
\]
\lau{06-09-2017: Should seals also have the option to be linear?}
\lau{13-09-2017: For now I have added linearity to the seals.}
The target language domains are all the non blue parts in the above. The source language domains are the black and blue parts in the above. Further
\begin{itemize}
\item $\linbnf$ defines domain $\Linear$
\item $\scbnf$ defines domain $\SealableCaps$
\item $\cbnf$ defines domain $\Caps$
\item $\Perm$ is defined as the set of permissions in Figure~\ref{fig:perm-hier}.
\end{itemize}

@@

\subsection{Syntax}
\label{sec:syntax}
The target machine is a simple capability machine with memory capabilities and sealed capabilities\footnote{In previous work, we used enter capabilities but for this the complexity introduced by mixing writable and executable memory is difficult to handle.} (inspired by CHERI). The syntax of the instructions of the target machine is defined as follows:
\[
\begin{array}{rcl}
n &\in & \nats \\
\trg{r} &\in &  \tRegName \\
\trg{\rn} &\defbnf &  \trg{r} \mid n \\
\trg{i} &\defbnf & \tfail \mid \thalt \mid \tjmp{\trg{r}} \mid \tjnz{\trg{r}}{\trg{\rn}} \mid \tisptr{\trg{r}}{\trg{r}} \mid \tgeta{\trg{r}}{\trg{r}} \mid \tgetb{\trg{r}}{\trg{r}} \mid \\
      & &  \tgete{\trg{r}}{\trg{r}}\mid \tgetp{\trg{r}}{\trg{r}} \mid \tgetlin{\trg{r}}{\trg{r}} \mid \tmove{\trg{r}}{\trg{\rn}} \mid \tstore{\trg{r}}{\trg{r}} \mid\\
      & &  \tload{\trg{r}}{\trg{r}} \mid \tcca{\trg{r}}{\trg{\rn}} \mid \trestrict{\trg{r}}{\trg{r}}{\trg{\rn}} \mid \tlt{\trg{r}}{\trg{r}}{\trg{r}} \mid \\
  & & \tplus{\trg{r}}{\trg{\rn}}{\trg{\rn}} \mid \tminus{\trg{r}}{\trg{\rn}}{\trg{\rn}} \mid \tsetatob{\trg{r}}{\trg{r}} \mid \txjmp{\trg{r}}{\trg{r}} \mid \tcseal{\trg{r}}{\trg{r}}{\trg{r}} \mid \\ 
      & &   \tsplit{\trg{r}}{\trg{r}}{\trg{r}} \mid\tsplice{\trg{r}}{\trg{r}}{\trg{r}} 
\end{array}
\]

@@

\[
  \isLinear{\cb} \defeq
  \begin{cases}
    \top & 
    \arraycolsep=0pt
    \begin{array}[t]{l}
      \cb = ((\_,\linear),\_,\_,\_) \tor\\
      \cb = \seal{\linear,\_,\_,\_} \tor\\
      \cb = \sealed{\cb',\_} \tand \isLinear{\cb'} 
    \end{array}\\
    \src{\top} & 
    \sourcecolor\left.
    \arraycolsep=0pt
    \begin{array}[t]{l}
      \cb = \stkptr{\_,\_,\_,\_}\\
      \cb = \retptrd(\_)
    \end{array}\right.\\
    \bot & \totherwise
  \end{cases}
\]

@@

\noindent\textbf{gettype}\\
In the definitions of the semantics below, we use a function $\encType{} : \Word \rightarrow \ints$. This is an encoding function for which the specific implementation does not matter. As the words of the two machines differ, we really need two functions which we call $\encType{}_{\var{src}}$ and $\encType{}_{\var{trg}}$. These two functions need to be related in the following way:
\lau{14-09-2017: Insert the relation. My impression is that the the words specific to the source machine should have the same "type" as their implementaion at the source level (I am not sure this is the case - maybe it is okay to just relate the values returned by the special words to the ones returned when used with the implementation). }
\begin{align*}
  \sem{\tisptr{r_1}{r_2}}(\Phi) = & \; \updPcAddr{}(\Phi\updReg{r_1}{\encType{\Phi(r_2)}})
\end{align*}

\noindent\textbf{geta, getb, gete, getp, and getl}\\
We assume function to encode and decode permissions as well as a function to encode and decode linearity. The functions are used implicitely when a permission or linearity is used in a place where they need to be a word.
\lau{14-09-2017: Write more about this (types and assumptions)}
\begin{align*}
  \sem{\tgeta{r_1}{r_2}}(\Phi) = & 
                                   \begin{cases}
                                     \updPcAddr{\Phi\updReg{r_1}{\aaddr}} & 
                                     \arraycolsep=0pt
                                     \begin{array}[t]{l}
                                       \Phi(r_2) = ((\_,\_),\_,\_,\aaddr) \\
                                       \quad \tor \Phi(r_2) = \seal{\_,\_,\_,\aaddr} \\
                                       \quad \src{\tor \Phi(r_2) = \stkptr{\_,\_,\_,\aaddr} } \\
                                     \end{array} \\
                                     \updPcAddr{\Phi\updReg{r_1}{-1}} & \totherwise
                                   \end{cases}
\end{align*}

\begin{align*}
  \sem{\tgetb{r_1}{r_2}}(\Phi) = & 
                                   \begin{cases}
                                     \updPcAddr{\Phi\updReg{r_1}{\baddr}} & 
                                     \arraycolsep=0pt
                                     \begin{array}[t]{l}
                                       \Phi(r_2) = ((\_,\_),\baddr,\_,\_) \\
                                       \quad \tor \Phi(r_2) = \seal{\_,\baddr,\_,\_} \\
                                       \quad \src{\tor \Phi(r_2) = \stkptr{\_,\baddr,\_,\_} } \\
                                     \end{array} \\
                                     \updPcAddr{\Phi\updReg{r_1}{-1}} & \totherwise
                                   \end{cases}
\end{align*}

\begin{align*}
  \sem{\tgete{r_1}{r_2}}(\Phi) = & 
                                   \begin{cases}
                                     \updPcAddr{\Phi\updReg{r_1}{\eaddr}} & 
                                     \arraycolsep=0pt
                                     \begin{array}[t]{l}
                                       \Phi(r_2) = ((\_,\_),\_,\eaddr,\_) \\
                                       \quad \tor \Phi(r_2) = \seal{\_,\_,\eaddr,\_} \\
                                       \quad \src{\tor \Phi(r_2) = \stkptr{\_,\_,\eaddr,\_} } \\
                                     \end{array} \\
                                     \updPcAddr{\Phi\updReg{r_1}{-1}} & \totherwise
                                   \end{cases}
\end{align*}
\lau{15-09-2017: Should gete and getb do something for seals?}

@@

\begin{align*}
  \sem{\tgetlin{r_1}{r_2}}(\Phi) = & 
                                   \begin{cases}
                                     \updPcAddr{\Phi\updReg{r_1}{\lin}} & 
                                     \arraycolsep=0pt
                                     \begin{array}[t]{l}
                                       \Phi(r_2) = ((\_,\lin),\_,\_,\_) \\
                                       \quad \tor \Phi(r_2) = \seal{\lin,\_,\_,\_} \\
                                       \quad \src{\tor \Phi(r_2) = \stkptr{\_,\_,\_,\_} \tand l = \linear}\\
                                     \end{array} \\
                                     \updPcAddr{\Phi\updReg{r_1}{-1}} & \totherwise
                                   \end{cases}
\end{align*}
\lau{15-09-2017: Should getl do something for seals?}
\lau{15-09-2017: Do we want to allow getl and getp to work on sealed capabilities?}

\noindent\textbf{move}\\
\begin{align*}
  \sem{\tmove{r_1}{r_2}}(\Phi) = & 
                              \begin{cases}
                                \updPcAddr{\Phi\updReg{r_1}{r_2}} & r_2 \in \ints \\
                                \updPcAddr{\Phi\updReg{r_1}{\Phi(r_2)}} & \nonLinear{\Phi(r_2)}\\
                                \updPcAddr{\Phi\updReg{r_2}{0}\updReg{r_1}{\Phi(r_2)}} & \totherwise 
                              \end{cases}
\end{align*}
Notice that in the case where we are moving a linear capability and $r_1 = r_2$ the order of the updates matter.

\noindent\textbf{store}\\
\begin{align*}
  \sem{\tstore{r_1}{r_2}}(\Phi) = &
                                    \begin{cases}
                                      \updPcAddr{}\left(
                                        \arraycolsep=0pt
                                        \begin{array}{rl}
                                          \Phi&\updReg{r_2}{w_2}\\
                                              &\update{\mem.a}{w}
                                        \end{array}
\right) & 
                                      \arraycolsep=0pt
                                      \begin{array}[t]{l}
                                        \Phi(r_1) = ((\perm,\lin),\baddr,\eaddr,\aaddr) \tand \\
                                        \perm \in \writeAllowed{} \tand\\
                                        \withinBounds{\Phi(r_1)} \tand \\
                                        w = \Phi(r_2) \tand \\
                                        w_2 = \linCons{w}
                                      \end{array}
                                      \\
                                      \sourcecolor\left.
                                      \updPcAddr{}\left(
                                      \arraycolsep=0pt
                                      \begin{array}{rl}
                                        \Phi&\update{\mem_\stk.a}{w}\\
                                            &\updReg{r_2}{w_2}
                                      \end{array}
                                      \right) \right.& 
                                      \sourcecolor\left.
                                      \arraycolsep=0pt
                                      \begin{array}[t]{l}
                                        \Phi(r_1) = \stkptr{\perm,\baddr,\eaddr,\aaddr} \tand \\
                                        \perm \in \writeAllowed{} \tand \\
                                        \withinBounds{\Phi(r_1)} \tand \\
                                        a \in \dom(\Phi.\ms_\stk) \tand \\
                                        w = \Phi(r_2) \\
                                        w_2 = \linCons{w}
                                      \end{array}\right.\\
                                      \failed & \totherwise
                                    \end{cases}
\end{align*}

\noindent\textbf{load}\\
\begin{align*}
  \sem{\tload{r_1}{r_2}}(\Phi) = & 
                                  \begin{cases}
                                    \updPcAddr{}\left(
                                      \arraycolsep=0pt
                                      \begin{array}{rl}
                                        \Phi&\updReg{r_1}{w}\\
                                            &\update{\mem.a}{w_2}
                                      \end{array}\right)& 
                                    \arraycolsep=0pt
                                    \begin{array}[t]{l}
                                      \Phi(r_2) = ((\perm,\lin),\baddr,\eaddr,\aaddr) \tand \\
                                      \perm \in \readAllowed{} \tand\\
                                      \withinBounds{\Phi(r_2)} \tand \\
                                      w = \Phi.\mem(a) \tand \\
                                      w_2 = \linCons{w}
                                    \end{array}
                                    \\
                                    \sourcecolor\left.
                                    \updPcAddr{}\left(
                                      \arraycolsep=0pt
                                      \begin{array}{rl}
                                        \Phi&\updReg{r_1}{w}\\
                                            & \update{\mem_\stk.a}{w_2}
                                      \end{array}\right)\right.
                                    & 
                                    \sourcecolor\left.
                                    \arraycolsep=0pt
                                    \begin{array}[t]{l}
                                      \Phi(r_2) = \stkptr{\perm,\baddr,\eaddr,\aaddr} \tand \\
                                      \perm \in \readAllowed{} \tand \\
                                      \withinBounds{\Phi(r_2)} \tand \\
                                      \aaddr \in \dom(\Phi.\ms_\stk) \tand \\
                                      w = \Phi.\ms_\stk(a) \tand \\
                                      w_2 = \linCons{w}
                                    \end{array}\right.
                                    \\
                                    \failed & \totherwise                                    
                                  \end{cases}
\end{align*}

\noindent\textbf{cca}\\
\emph{Change Current Address}
\lau{15-09-2017: This is the old lea. I changed the name, so it actually reflects what the instruction does. We might also want to add a real lea at some point.}
\begin{align*}
  \sem{\tcca{r_1}{r_2}}(\Phi) = & 
                                  \begin{cases}
                                    \updPcAddr{\Phi\updReg{r_1}{c}} &  
                                    \arraycolsep=0pt
                                    \begin{array}[t]{l}
                                      \text{either $n = r_2$ or $n = \Phi.\reg(r_2)$}\\
                                      \quad\text{ and in either case $n \in \ints$ and} \\
                                      \Phi(r_1) = ((\perm,\lin),\baddr,\eaddr,\aaddr) \tand \\
                                      \perm \neq \enter \tand \\
                                      c = ((\perm,\lin),\baddr,\eaddr,\aaddr + n)
                                    \end{array}
                                    \\
                                    \updPcAddr{\Phi\updReg{r_1}{c}} &  
                                    \arraycolsep=0pt
                                    \begin{array}[t]{l}
                                      \text{either $n = r_2$ or $n = \Phi.\reg(r_2)$}\\
                                      \quad\text{ and in either case $n \in \ints$ and} \\
                                      \Phi(r_1) = \seal{\lin,\sigma_\baddr,\sigma_\eaddr,\sigma} \tand \\
                                      c = \seal{\lin,\sigma_\baddr,\sigma_\eaddr,\sigma + n}
                                    \end{array}
                                    \\
                                    \src{\updPcAddr{\Phi\updReg{r_1}{c}}} &  
                                    \sourcecolor\left.
                                    \arraycolsep=0pt
                                    \begin{array}[t]{l}
                                      \text{either $n = r_2$ or $n = \Phi(r_2)$}\\
                                      \quad\text{ and in either case $n \in \ints$ and} \\
                                      \Phi(r_1) = \stkptr{\perm,\baddr,\eaddr,\aaddr} \tand \\
                                      \perm \neq \enter \tand \\
                                      c = \stkptr{\perm,\baddr,\eaddr,\aaddr + n}
                                    \end{array}\right.
                                    \\
                                    \failed & \totherwise
                                  \end{cases}
\end{align*}

@@

@@

                                  \Phi(r_3) = ((\perm,\lin),\baddr_3,\eaddr_3,\_) \tand \\

@@

                                  \Phi(r_2) = \seal{\lin,\sigma_{\baddr,2},\sigma_{\eaddr,2},\sigma} \tand \\

>>>>

@@

% Four arguments
\newcommand{\ssubseg}[4]{\fourinstr{\src{subseg}}{#1}{#2}{#3}{#4}}
\newcommand{\ssplit}[4]{\fourinstr{\src{split}}{#1}{#2}{#3}{#4}}

@@

% Four arguments
\newcommand{\tsubseg}[4]{\fourinstr{\trg{subseg}}{#1}{#2}{#3}{#4}}
\newcommand{\tsplit}[4]{\fourinstr{\trg{split}}{#1}{#2}{#3}{#4}}

@@

\dominique{7-9-2017: Should we keep local/global and linear/non-linear?  I was thinking just the latter would suffice...}
\lau{07-09-2017: The local/global mechanism still has some value as it now can be used for the purpose of arguments that cannot be stored by the receiver (if you are willing to ``pay'' the clearing cost). I do, however, not think that it is strictly necessary here to make things work.}
\lau{12-09-2017: I will remove the local capabilities today for the reasons we have discussed (they add to the complexity and they seem to be of little to no use without a clear memory instruction).}
Source language domains:
\[
  \begin{array}{rrcl}
   \addrbnf,\basebnf \in & \Addr & \defeq & \nats \\
    \sealbasebnf, \sigma \in & \Seal & \defeq & \nats \\
    &\Word & \defeq & \ints \uplus \SealableCaps\\
    \permbnf \in& \Perm & \defbnf & \dots \\
    &\linbnf & \defbnf & \linear \mid \normal \\
    &\aendbnf & \in & \Addr \uplus \{\infty \} \\
    &\sealendbnf & \in & \Seal \uplus \{\infty \} \\
    &\scbnf &\defbnf & ((\permbnf,\linbnf),\basebnf,\aendbnf,\addrbnf) \mid \seal{\linbnf,\sealbasebnf,\sealendbnf}\\
    & & & {\sourcecolor{} \mid \stkptr{\permbnf,\basebnf,\aendbnf,\addrbnf}}\\ 
    &\cbnf & \defbnf &  \scbnf \mid \sealed{\scbnf,\sigma}{\sourcecolor{} \; \mid \retptrd(\sigma) \mid \retptrc(\sigma)}\\ 
    &\RegName & \defbnf &\pcreg \mid \rret \mid \rstk \mid \rdata \mid \dots \\
    &\Reg & \defeq & \RegName \fun \Word\\
    &\Mem & \defeq & \Addr \fun \Word \\
    &\MemSeg & \defeq & \Addr \parfun \Word \\
    {\sourcecolor{} \sstkframe \in} & {\sourcecolor{} \StkFrame} & {\sourcecolor{} \defeq} & {\sourcecolor{} \Caps \times \MemSeg \times \Addr}\\
    \src{\sstk \in}& \src{ \Stack} & \src{ \defeq} & \src{ \StkFrame^*} \\
    \Phi \in & \ExecConf & \defeq & \Mem \times \Reg {\sourcecolor{} \; \times \; \Stack \times \MemSeg }\\
    &\Conf & \defeq & \ExecConf \uplus \{\failed\} \uplus (\{\halted\} \times \Mem)
  \end{array}
\]
\lau{06-09-2017: Should seals also have the option to be linear?}
\lau{13-09-2017: For now I have added linearity to the seals.}
The target language domains are all the non blue parts in the above. The source language domains are the black and blue parts in the above. Further
\begin{itemize}
\item $\linbnf$ defines domain $\Linear$
\item $\scbnf$ defines domain $\SealableCaps$
\item $\cbnf$ defines domain $\Caps$
\item $\Perm$ is defined as the set of permissions in Figure~\ref{fig:perm-hier}.
\end{itemize}

@@

\subsection{Syntax}
\label{sec:syntax}
The target machine is a simple capability machine with memory capabilities and sealed capabilities\footnote{In previous work, we used enter capabilities but for this the complexity introduced by mixing writable and executable memory is difficult to handle.} (inspired by CHERI). The syntax of the instructions of the target machine is defined as follows:
\[
\begin{array}{rcl}
n &\in & \nats \\
\trg{r} &\in &  \tRegName \\
\trg{\rn} &\defbnf &  \trg{r} \mid n \\
\trg{i} &\defbnf & \tfail \mid \thalt \mid \tjmp{\trg{r}} \mid \tjnz{\trg{r}}{\trg{\rn}} \mid \tisptr{\trg{r}}{\trg{r}} \mid \tgeta{\trg{r}}{\trg{r}} \mid \tgetb{\trg{r}}{\trg{r}} \mid \\
      & &  \tgete{\trg{r}}{\trg{r}}\mid \tgetp{\trg{r}}{\trg{r}} \mid \tgetlin{\trg{r}}{\trg{r}} \mid \tmove{\trg{r}}{\trg{\rn}} \mid \tstore{\trg{r}}{\trg{r}} \mid\\
      & &  \tload{\trg{r}}{\trg{r}} \mid \tcca{\trg{r}}{\trg{\rn}} \mid \trestrict{\trg{r}}{\trg{r}}{\trg{\rn}} \mid \tsubseg{\trg{r}}{\trg{r}}{\trg{\rn}}{\trg{\rn}} \mid \tlt{\trg{r}}{\trg{r}}{\trg{r}} \mid \\
  & & \tplus{\trg{r}}{\trg{\rn}}{\trg{\rn}} \mid \tminus{\trg{r}}{\trg{\rn}}{\trg{\rn}} \mid \tsetatob{\trg{r}}{\trg{r}} \mid \txjmp{\trg{r}}{\trg{r}} \mid \tcseal{\trg{r}}{\trg{r}}{\trg{r}} \mid \\ 
      & &   \tsplit{\trg{r}}{\trg{r}}{\trg{r}} \mid\tsplice{\trg{r}}{\trg{r}}{\trg{r}} 
\end{array}
\]

@@

\[
  \isLinear{\cb} \defeq
  \begin{cases}
    \top & 
    \arraycolsep=0pt
    \begin{array}[t]{l}
      \cb = ((\_,\linear),\_,\_,\_) \tor\\
      \cb = \seal{\linear,\_,\_} \tor\\
      \cb = \sealed{\cb',\_} \tand \isLinear{\cb'} 
    \end{array}\\
    \src{\top} & 
    \sourcecolor\left.
    \arraycolsep=0pt
    \begin{array}[t]{l}
      \cb = \stkptr{\_,\_,\_,\_}\\
      \cb = \retptrd(\_)
    \end{array}\right.\\
    \bot & \totherwise
  \end{cases}
\]

@@

\noindent\textbf{gettype}\\
In the definitions of the semantics below, we use a function $\encType{} : \Word \rightarrow \ints$. This is an encoding function for which the specific implementation does not matter. As the words of the two machines differ, we really need two functions which we call $\encType{}_{\var{src}}$ and $\encType{}_{\var{trg}}$. These two functions need to be related in the following way:
\lau{14-09-2017: Insert the relation. My impression is that the the words specific to the source machine should have the same "type" as their implementaion at the source level (I am not sure this is the case - maybe it is okay to just relate the values returned by the special words to the ones returned when used with the implementation). }
\dominique{20-9-2017: I also think gettype for source-specific words should produce the same result than gettype for their implementation.}
\begin{align*}
  \sem{\tisptr{r_1}{r_2}}(\Phi) = & \; \updPcAddr{}(\Phi\updReg{r_1}{\encType{\Phi(r_2)}})
\end{align*}

\noindent\textbf{geta, getb, gete, getp, and getl}\\
We assume function to encode and decode permissions as well as a function to encode and decode linearity. The functions are used implicitely when a permission or linearity is used in a place where they need to be a word.
\lau{14-09-2017: Write more about this (types and assumptions)}
\begin{align*}
  \sem{\tgeta{r_1}{r_2}}(\Phi) = & 
                                   \begin{cases}
                                     \updPcAddr{\Phi\updReg{r_1}{\aaddr}} & 
                                     \arraycolsep=0pt
                                     \begin{array}[t]{l}
                                       \Phi(r_2) = ((\_,\_),\_,\_,\aaddr) \\
                                       \quad \src{\tor \Phi(r_2) = \stkptr{\_,\_,\_,\aaddr} } \\
                                     \end{array} \\
                                     \updPcAddr{\Phi\updReg{r_1}{-1}} & \totherwise
                                   \end{cases}
\end{align*}

\begin{align*}
  \sem{\tgetb{r_1}{r_2}}(\Phi) = & 
                                   \begin{cases}
                                     \updPcAddr{\Phi\updReg{r_1}{\baddr}} & 
                                     \arraycolsep=0pt
                                     \begin{array}[t]{l}
                                       \Phi(r_2) = ((\_,\_),\baddr,\_,\_) \\
                                       \quad \src{\tor \Phi(r_2) = \stkptr{\_,\baddr,\_,\_} } \\
                                     \end{array} \\
                                     \updPcAddr{\Phi\updReg{r_1}{-1}} & \totherwise
                                   \end{cases}
\end{align*}

\begin{align*}
  \sem{\tgete{r_1}{r_2}}(\Phi) = & 
                                   \begin{cases}
                                     \updPcAddr{\Phi\updReg{r_1}{\eaddr}} & 
                                     \arraycolsep=0pt
                                     \begin{array}[t]{l}
                                       \Phi(r_2) = ((\_,\_),\_,\eaddr,\_) \\
                                       \quad \src{\tor \Phi(r_2) = \stkptr{\_,\_,\eaddr,\_} } \\
                                     \end{array} \\
                                     \updPcAddr{\Phi\updReg{r_1}{-1}} & \totherwise
                                   \end{cases}
\end{align*}
\lau{15-09-2017: Should gete and getb do something for seals?}

@@

\begin{align*}
  \sem{\tgetlin{r_1}{r_2}}(\Phi) = & 
                                   \begin{cases}
                                     \updPcAddr{\Phi\updReg{r_1}{\lin}} & 
                                     \arraycolsep=0pt
                                     \begin{array}[t]{l}
                                       \Phi(r_2) = ((\_,\lin),\_,\_,\_) \\
                                       \quad \src{\tor \Phi(r_2) = \stkptr{\_,\_,\_,\_} } \tand l = \linear\\
                                     \end{array} \\
                                     \updPcAddr{\Phi\updReg{r_1}{-1}} & \totherwise
                                   \end{cases}
\end{align*}
\lau{15-09-2017: Should getl do something for seals?}
\lau{15-09-2017: Do we want to allow getl and getp to work on sealed capabilities?}

\noindent\textbf{move}\\
\begin{align*}
  \sem{\tmove{r_1}{r_2}}(\Phi) = & 
                              \begin{cases}
                                \updPcAddr{\Phi\updReg{r_1}{\Phi(r_2)}} & \nonLinear{\Phi(r_2)}\\
                                \updPcAddr{\Phi\updReg{r_2}{0}\updReg{r_1}{\Phi(r_2)}} & \totherwise 
                              \end{cases}
\end{align*}
Notice that in the case where we are moving a linear capability and $r_1 = r_2$ the order of the updates matter.

\noindent\textbf{store}\\
\begin{align*}
  \sem{\tstore{r_1}{r_2}}(\Phi) = &
                                    \begin{cases}
                                      \updPcAddr{}\left(
                                        \arraycolsep=0pt
                                        \begin{array}{rl}
                                          \Phi&\updReg{r_2}{w_2}\\
                                              &\update{\mem.a}{w}
                                        \end{array}
\right) & 
                                      \arraycolsep=0pt
                                      \begin{array}[t]{l}
                                        \Phi(r_1) = ((\perm,\lin),\baddr,\eaddr,\aaddr) \tand \\
                                        \perm \in \writeAllowed{} \tand\\
                                        \withinBounds{\Phi(r_1)} \tand \\
                                        w = \Phi(r_2) \tand \\
                                        w_2 = \linCons{w}
                                      \end{array}
                                      \\
                                      \sourcecolor\left.
                                      \updPcAddr{}\left(
                                      \arraycolsep=0pt
                                      \begin{array}{rl}
                                        \Phi&\update{\mem.a}{w}\\
                                            &\updReg{r_2}{w_2}
                                      \end{array}
                                      \right) \right.& 
                                      \sourcecolor\left.
                                      \arraycolsep=0pt
                                      \begin{array}[t]{l}
                                        \Phi(r_1) = \stkptr{\perm,\baddr,\eaddr,\aaddr} \tand \\
                                        \perm \in \writeAllowed{} \tand \\
                                        \withinBounds{\Phi(r_1)} \tand \\
                                        a \in \dom(\Phi.\ms_\stk) \tand \\
                                        w = \Phi(r_2) \\
                                        w_2 = \linCons{w}
                                      \end{array}\right.\\
                                      \failed & \totherwise
                                    \end{cases}
\end{align*}

\noindent\textbf{load}\\
\begin{align*}
  \sem{\tload{r_1}{r_2}}(\Phi) = & 
                                  \begin{cases}
                                    \updPcAddr{}\left(
                                      \arraycolsep=0pt
                                      \begin{array}{rl}
                                        \Phi&\updReg{r_1}{w}\\
                                            &\update{\mem.a}{w_2}
                                      \end{array}\right)& 
                                    \arraycolsep=0pt
                                    \begin{array}[t]{l}
                                      \Phi(r_2) = ((\perm,\lin),\baddr,\eaddr,\aaddr) \tand \\
                                      \perm \in \readAllowed{} \tand\\
                                      \withinBounds{\Phi(r_2)} \tand \\
                                      w = \Phi.\mem(a) \tand \\
                                      w_2 = \linCons{w}
                                    \end{array}
                                    \\
                                    \sourcecolor\left.
                                    \updPcAddr{}\left(
                                      \arraycolsep=0pt
                                      \begin{array}{rl}
                                        \Phi&\updReg{r_1}{w}\\
                                            & \update{\mem_\stk.a}{w_2}
                                      \end{array}\right)\right.
                                    & 
                                    \sourcecolor\left.
                                    \arraycolsep=0pt
                                    \begin{array}[t]{l}
                                      \Phi(r_2) = \stkptr{\perm,\baddr,\eaddr,\aaddr} \tand \\
                                      \perm \in \readAllowed{} \tand \\
                                      \withinBounds{\Phi(r_2)} \tand \\
                                      \in \dom(\Phi.\ms_\stk) \tand \\
                                      w = \Phi.\ms_\stk(a) \tand \\
                                      w_2 = \linCons{w}
                                    \end{array}\right.
                                    \\
                                    \failed & \totherwise                                    
                                  \end{cases}
\end{align*}

\noindent\textbf{cca}\\
\emph{Change Current Address}
\lau{15-09-2017: This is the old lea. I changed the name, so it actually reflects what the instruction does. We might also want to add a real lea at some point.}
\begin{align*}
  \sem{\tcca{r_1}{r_2}}(\Phi) = & 
                                  \begin{cases}
                                    \updPcAddr{\Phi\updReg{r_1}{c}} &  
                                    \arraycolsep=0pt
                                    \begin{array}[t]{l}
                                      \text{either $n = r_2$ or $n = \Phi.\reg(r_2)$}\\
                                      \quad\text{ and in either case $n \in \ints$ and} \\
                                      \Phi(r_1) = ((\perm,\lin),\baddr,\eaddr,\aaddr) \tand \\
                                      \perm \neq \enter \tand \\
                                      c = ((\perm,\lin),\baddr,\eaddr,\aaddr + n)
                                    \end{array}
                                    \\
                                    \src{\updPcAddr{\Phi\updReg{r_1}{c}}} &  
                                    \sourcecolor\left.
                                    \arraycolsep=0pt
                                    \begin{array}[t]{l}
                                      \text{either $n = r_2$ or $n = \Phi(r_2)$}\\
                                      \quad\text{ and in either case $n \in \ints$ and} \\
                                      \Phi(r_1) = \stkptr{\perm,\baddr,\eaddr,\aaddr} \tand \\
                                      \perm \neq \enter \tand \\
                                      c = \stkptr{\perm,\baddr,\eaddr,\aaddr + n}
                                    \end{array}\right.
                                    \\
                                    \failed & \totherwise
                                  \end{cases}
\end{align*}

@@

\noindent\textbf{subseg}\\
\lau{19-09-2017: TODO: I am not sure the way the end points are handled makes entirely sense. Have a look at it (maybe make a function that computes the end point?)}
  \sem{\tsubseg{r_1}{r_2}{r_3}{r_4}}(\Phi) = &
                                                 \Phi\updReg{r_2}{w_2}\updReg{r_1}{c} &
                                                   \text{for $i \in \{3, 4\}$ } \\
                                                   \quad \text{either $n_i = r_i$ or $n_i = \Phi.\reg(r_i)$}\\
                                                   \quad \text{and in either case $n_i \in \ints$ and} \\
                                                   \Phi(r_2) = ((\perm,\lin),\baddr,\eaddr,\aaddr) \tand \\
                                                   \baddr \leq n_3 \tand \\
                                                   (n_2 \leq \eaddr \tand n_2 \in \nats) \Rightarrow \eaddr' = n_2 \tand \\
                                                   (n_2 = -42 \tand \eaddr = \infty) \Rightarrow \eaddr' = \infty \tand\\
                                                   \perm \neq \enter \tand\\
                                                   c = ((\perm,\lin),n_1,\eaddr',\aaddr)\\
                                                   w_2 = \linCons{\Phi(r_2)}
                                                 \end{array}
\\
                                                 \Phi\updReg{r_2}{w_2}\updReg{r_1}{c} &
                                                   \text{for $i \in \{3, 4\}$ } \\
                                                   \quad \text{either $n_i = r_i$ or $n_i = \Phi.\reg(r_i)$}\\
                                                   \quad \text{and in either case $n_i \in \ints$ and} \\
                                                   \Phi(r_2) = \seal{\lin,\sigma_\baddr,\sigma_\eaddr} \tand \\
                                                   \baddr \leq n_3 \tand \\
                                                   (n_2 \leq \eaddr \tand n_2 \in \nats) \Rightarrow \eaddr' = n_2 \tand \\
                                                   (n_2 = -42 \tand \eaddr = \infty) \Rightarrow \eaddr' = \infty \tand\\
                                                   \perm \neq \enter \tand\\
                                                   c = \seal{\lin,n_1,\eaddr'}\\
                                                   w_2 = \linCons{\Phi(r_2)}
                                                 \end{array}
\\
                                                 \src{\Phi\updReg{r_2}{0}\updReg{r_1}{c}} &
                                                   \text{for $i \in \{3, 4\}$ } \\
                                                   \quad \text{either $n_i = r_i$ or $n_i = \Phi.\reg(r_i)$}\\
                                                   \quad \text{and in either case $n_i \in \ints$ and} \\
                                                   \Phi(r_2) = \stkptr{\perm,\baddr,\eaddr,\aaddr} \tand \\
                                                   \baddr \leq n_3 \tand \\
                                                   (n_2 \leq \eaddr \tand n_2 \in \nats) \Rightarrow \eaddr' = n_2 \tand\\
                                                   (n_2 = -42 \tand \eaddr = \infty) \Rightarrow \eaddr' = \infty \tand\\
                                                   \perm \neq \enter \tand\\
                                                   c = \stkptr{\perm,n_1,\eaddr',\aaddr}\\
                                                 \end{array}\right.\\

@@

                                  \Phi(r_3) = ((\perm,\lin),\baddr_3,\eaddr_3,\aaddr_3) \tand \\

@@

                                  \Phi(r_2) = \seal{\lin,\sigma_{\baddr,2},\sigma_{\eaddr,2}} \tand \\
